Fission product decay chain equilibria in \zirconia

### 2019-02-11 manuscript.tex
Update intro

%\begin{abstract}

%\end{abstract}

@@

Stress-corrosion cracking (SCC) in nuclear fuel pins is an issue related to early loss of structural integrity of fuel assemblies in light water reactors (LWRs). In particular, the phenomenon of pellet-cladding interaction (PCI) in combination with SCC can lead to failures where the cladding is breached, exposing fuel to the coolant \cite{Cox1990}.     \\

Nuclear fuel claddings have unique materials challenges owing to the highly active environment in which they operate and the creation of unstable isotopes. Corrosive species in the pin such as iodine, produced directly as a result of uranium fission are known to play a role in SCC [REF]. However, iodine nuclei produced from fission are unstable. Fission of uranium produces iodine precursors, mainly unstable isotopes of tellurium. Both iodine and tellurium are relatively common fission products, with combined independent yields from thermal fission of U$_{235}$ above 5\% \cite{kennett1956mass, iodine129fissionyield, imanishi1976independent, iaeafissionyield, iodine132, amiel1975odd}. In addition, xenon and caesium are also common fission products, as shown in Figure \ref{fig:fissionyield}, with fission product masses firmly within the heavy section of the bi-modal fission product distribution curve.    \\

Previous work on defect equilibria in \zirconia\ considered the internal oxide layer's effectiveness as a barrier to iodine \cite{kenichiodine2018}. It was found that the tetragonal phase of \zirconia\ is a better barrier to iodine ingress than monoclinic \zirconia , especially at higher oxygen partial pressures. It is also established that tetragonal \zirconia\ will always be present on the inner surface of the cladding in significant quantities because it is self-stabilised by the stresses imposed as the oxide grows into the zirconium metal, in addition to compressive residual stresses induced by radiation damage. The iodine defect study, however, only informs us about one part of the SCC process. For a more holistic understanding, the life cycle of the iodine must also be taken into account.     \\

@@

Nuclei produced during fission are typically neutron-rich, resulting in decay modes such as $\beta-$ or neutron emission. In the case of tellurium nuclei, most decay into iodine, which then decays into xenon with different half-lives depending on the isotope. The decay chain continues with xenon nuclei decaying into caesium, many isotopes of which have half lives measured in years. At this point, conventional power reactor fuel is retired long before a significant proportion of the caesium decays into barium. For this reason we only consider the elements tellurium through caesium in this study. \\

The majority of thermal fission events occur in the outer rim of a fuel pellet. Given a fission product penetration depth of up to 8 $\mu$m in \zirconia\ \cite{degueldre2001behaviour}, a large proportion of fission products implant within the oxide. \\

@@

Brouwer diagrams for caesium defects are shown in Figure \ref{figure:caesiumbrouwer}. Caesium defects were mostly unaffected by a change in oxygen pressure, with dominant defects only changing from \ch{Cs_{O}^{**}} to \ch{Cs_{Zr}^{'''}} at an oxygen pressure of $10^{-30.5}$ atm, regardless of the caesium concentration. However, because additional charge compensation for caesium defects is necessary at higher caesium concentrations, we see an increase in the concentration of \ch{V_{O}^{**}} to compensate for the negative \ch{Cs_{Zr}^{'''}} defects.

@@

As fission products proceed down the decay chain from Te to Cs, the defects they produce in \zirconia\ have progressively larger volumes. One of the largest changes in defect volume occurs when \ch{Te_{Zr}^{'''}} decays and produces \ch{I_{Zr}^{'''}} (+27.14 \r{A}$^{3}$). This may then decay and produce the defect \ch{Xe_{Zr}^{''''}}, which leads to an even larger change in defect volume (+28.43 \r{A}$^{3}$).

@@

%\lhead{\emph{References}} % Change the page header to say "References"
\bibliographystyle{unsrt}
\bibliography{Mendeley, Bibliography} % The references (bibliography) information are stored in the file named "Bibliography.bib"

@@

\begin{table}[htp]
\centering
\caption{Defect volumes and formation energies of the dominant defect types in tetragonal \zirconia .}
\label{defect_volumes}
\begin{tabular}{ccc}
\hline
Defect type & Defect volume (\r{A}$^{3}$) & Formation energy at 1500 K (eV) \\ \hline
\ch{V_{O}^{**}} & -40.45 & -0.92 \\ %7440 \\ %362 \\
\ch{V_{Zr}^{''''}} & 88.93 & 7.04 \\ %9072 \\ %8347 \\
%\ch{V_{O}^{*}} & -21.14 \\ %3522 \\
%\ch{V_{O}^{x}} & -2.96 \\ %415 \\
\ch{Te_{O}^{***}} & -18.83 & 1.07 \\ %8097 \\ %6605 \\
\ch{Te_{O}^{**}} & 2.74 & 2.24 \\ %5940 \\ %3752 \\
\ch{Te_{O}^{*}} & 24.61 & 5.10 \\ %5090 \\ %7451 \\
\ch{Te_{Zr}^{''''}} & 83.01 & 8.85 \\ %6289 \\ %8247 \\
\ch{Te_{Zr}^{'''}} & 59.93 & 5.97 \\ %0934 \\ %2232 \\
\ch{I_{O}^{***}} & -18.59 & -2.68 \\ %6196 \\ %0449 \\
\ch{I_{O}^{*}} & 29.90 & 0.10 \\ %8003 \\ %6407 \\
\ch{I_{Zr}^{'''}} & 87.07 & 4.17 \\ %5444 \\ %912 \\
\ch{Xe_{O}^{**}} & 9.02 & -1.18 \\ %8234 \\ %3892 \\
\ch{Xe_{O}^{*}} & 32.07 & 1.84 \\ %7508 \\ %5247 \\
\ch{Xe_{Zr}^{''''}} & 115.50 & 6.72 \\ %2321 \\ %4415 \\
\ch{Xe_{Zr}^{'''}} & 91.65 & 4.97 \\ %1595 \\ %578 \\
\ch{Cs_{O}^{**}} & 4.38 & 4.17 \\ %0571 \\ %1495 \\
\ch{Cs_{Zr}^{'''}} & 90.37 & 5.97 \\ \hline %0931 \\ \hline %6229 \\ \hline
\end{tabular}
\end{table}

>>>>

\begin{abstract}

As fission products proceed down the decay chain from Te to Cs, the defects they produce in \zirconia\ have progressively larger volumes. At the oxygen site following fission product implantation, Te in a 0 oxidation state will form the defect \ch{Te_{O}^{**}} as the dominant anion site defect, with a defect volume relative to the non-defective crystal of +2.74 \r{A}$^{3}$. These Te nuclei  One of the largest changes in defect volume occurs when \ch{Te_{Zr}^{'''}} decays and produces \ch{I_{Zr}^{'''}} (+27.14 \r{A}$^{3}$). This may then decay and produce the defect \ch{Xe_{Zr}^{''''}}, which leads to an even larger change in defect volume (+28.43 \r{A}$^{3}$).

\end{abstract}

@@

% the time delay before failure in pci related failures is an important thing to talk about. cite that paper which you bookmarked. Talk about how the

Stress-corrosion cracking (SCC) in nuclear fuel pins is an issue related to early loss of structural integrity of fuel assemblies in light water reactors (LWRs). In particular, the phenomenon of pellet-cladding interaction (PCI) in combination with SCC can lead to failures where the cladding is breached, exposing fuel to the coolant \cite{bcoxpelletclad1990}.     \\

Nuclear fuel claddings have unique materials challenges owing to the highly active environment in which they operate and the creation of unstable isotopes. Corrosive species in the pin such as iodine, produced directly as a result of uranium fission are known to play a role in SCC \cite{Sidky1998, Fregonese1998, iodinezrmetal}. However, iodine nuclei produced from fission are unstable. Fission of uranium produces iodine precursors, mainly unstable isotopes of tellurium. Both iodine and tellurium are relatively common fission products, with combined independent yields from thermal fission of U$_{235}$ above 5\% \cite{kennett1956mass, iodine129fissionyield, imanishi1976independent, iaeafissionyield, iodine132, amiel1975odd}. In addition, xenon and caesium are also common fission products, as shown in Figure \ref{fig:fissionyield}, with fission product masses firmly within the heavy nuclide peak of uranium and plutonium fission product distribution curves.    \\

Previous work on defect equilibria in \zirconia\ considered the internal oxide layer's effectiveness as a barrier to iodine \cite{kenichiodine2018}. It was found that the tetragonal phase of \zirconia\ is a better barrier to iodine ingress than monoclinic \zirconia , especially at higher oxygen partial pressures. It is also established that tetragonal \zirconia\ will always be present on the inner surface of the cladding in significant quantities because it is self-stabilised by the stresses imposed as the oxide grows into the zirconium metal, in addition to compressive residual stresses induced by radiation damage \cite{causey2005review}. The iodine defect study, however, only informs us about one part of the SCC process. For a more holistic understanding, the life cycle of the iodine must also be taken into account.     \\

@@

Nuclei produced during fission are typically neutron-rich, resulting in decay modes such as $\beta-$ or neutron emission. In the case of tellurium nuclei, most decay into iodine, which then decays into xenon with different half-lives depending on the isotope. Several of these isotopes' fission yield and decay data are shown in Table \ref{table:decaydata}. The decay chain continues with xenon nuclei decaying into caesium, many isotopes of which have half lives measured in years. At this point, conventional power reactor fuel is retired long before a significant proportion of the caesium decays into barium. For this reason we only consider the elements tellurium through caesium in this study. \\

The majority of thermal fission events occur in the outer rim of a fuel pellet. Given a fission product penetration depth of up to 8 $\mu$m in \zirconia\ \cite{degueldre2001behaviour}, a large proportion of fission products implant within the oxide. It has also been observed in PCI failures, that there is a time delay on the order of 10 minutes between the power ramping of the reactor and the subsequent breach of the cladding \cite{bergenlid1980experimental, wood1983effects, pankaskie1981mechanistic}. This time delay is in line with the half lives of many Te and I isotopes (Table \ref{table:decaydata}) such that significant proportions of these isotopes will decay during this time delay, hinting that these phenomena could be related. \\
%In addition, it is known that irradiated cladding containing fresh fuel does not crack when power ramped  \cite{wood1983effects}. \\

@@

Brouwer diagrams for caesium defects are shown in Figure \ref{figure:caesiumbrouwer}. Caesium defects were mostly unaffected by a change in oxygen pressure, with dominant defects only changing from \ch{Cs_{O}^{**}} to \ch{Cs_{Zr}^{'''}} at an oxygen pressure of $10^{-30.5}$ atm, regardless of the caesium concentration. However, because additional charge compensation for caesium defects is necessary at higher caesium concentrations, we see an increase in the concentration of \ch{V_{O}^{**}} to compensate for the negative \ch{Cs_{Zr}^{'''}} defects. \\

The defect volumes of \ch{Cs_{O}^{**}} and \ch{Cs_{Zr}^{'''}} were 4.38 and 90.37 \r{A}$^{3}$ respectively, less than the Xe defects on the same site. This is consistent with the fact that Cs has a much smaller first ionisation energy than Xe, therefore presenting a smaller energy barrier to reducing it's size on either site to improve lattice fit. Because of this, we do not expect further stresses being imposed on the lattice after Xe decays to Cs. We can therefore consider only Te, I and Xe to be significant for this mechanism, as the only contribution to lattice stresses due to Cs would have occurred by direct implantation after fission.

%We can therefore not consider Cs in this process, narrowing it down to Te, I, and Xe.

% healing radiation damage occurs much faster than the decay of isotopes.

@@

As fission products proceed down the decay chain from Te to Xe, the defects they produce in \zirconia\ have progressively larger volumes. One of the largest changes in defect volume occurs when \ch{Te_{Zr}^{'''}} decays and produces \ch{I_{Zr}^{'''}} (+27.14 \r{A}$^{3}$). This may then decay and produce the defect \ch{Xe_{Zr}^{''''}}, which leads to an even larger change in defect volume (+28.43 \r{A}$^{3}$). This process may be a contributing factor to fuel cladding failures observed after a power ramp, with reported time delays between ramp and failure aligning with the decay rate of Te and I isotopes.

@@

%\lhead{\emph{References}} % Change the page header to say "References"
\bibliographystyle{unsrt}
\bibliography{Mendeley} % The references (bibliography) information are stored in the file named "Bibliography.bib"

@@

\begin{table}[htp]
\centering
\caption{Defect volumes and formation energies of the dominant defect types in tetragonal \zirconia .}
\label{defect_volumes}
\begin{tabular}{ccc}
\hline
Defect type & \begin{tabular}[c]{@{}c@{}}Defect volume relative\\ to perfect crystal (\r{A}$^{3}$)\end{tabular} & \begin{tabular}[c]{@{}c@{}}Formation energy \\ at 1500 K (eV)\end{tabular} \\ \hline
\ch{V_{O}^{**}} & -40.45 & -0.92 \\ %7440 \\ %362 \\
\ch{V_{Zr}^{''''}} & 88.93 & 7.04 \\ %9072 \\ %8347 \\
%\ch{V_{O}^{*}} & -21.14 \\ %3522 \\
%\ch{V_{O}^{x}} & -2.96 \\ %415 \\
\ch{Te_{O}^{***}} & -18.83 & 1.07 \\ %8097 \\ %6605 \\
\ch{Te_{O}^{**}} & 2.74 & 2.24 \\ %5940 \\ %3752 \\
\ch{Te_{O}^{*}} & 24.61 & 5.10 \\ %5090 \\ %7451 \\
\ch{Te_{Zr}^{''''}} & 83.01 & 8.85 \\ %6289 \\ %8247 \\
\ch{Te_{Zr}^{'''}} & 59.93 & 5.97 \\ %0934 \\ %2232 \\
\ch{I_{O}^{***}} & -18.59 & -2.68 \\ %6196 \\ %0449 \\
\ch{I_{O}^{*}} & 29.90 & 0.10 \\ %8003 \\ %6407 \\
\ch{I_{Zr}^{'''}} & 87.07 & 4.17 \\ %5444 \\ %912 \\
\ch{Xe_{O}^{**}} & 9.02 & -1.18 \\ %8234 \\ %3892 \\
\ch{Xe_{O}^{*}} & 32.07 & 1.84 \\ %7508 \\ %5247 \\
\ch{Xe_{Zr}^{''''}} & 115.50 & 6.72 \\ %2321 \\ %4415 \\
\ch{Xe_{Zr}^{'''}} & 91.65 & 4.97 \\ %1595 \\ %578 \\
\ch{Cs_{O}^{**}} & 4.38 & 4.17 \\ %0571 \\ %1495 \\
\ch{Cs_{Zr}^{'''}} & 90.37 & 5.97 \\ \hline %0931 \\ \hline %6229 \\ \hline
\end{tabular}
\end{table}

\begin{table}
\caption{Independent fission product yields and half-lives for the major iodine isotopes and precursors in a thermal neutron reactor. Yields from \cite{iodinefissionyields, iodine132}. All isotopes undergo single $\beta$- decay. Metastable states are included.}  \label{table:decaydata}
\begin{center}
\begin{tabular}{c c c}
\hline
Isotope & Independent Yield (\%) & Half-life \\
\hline
\texorpdfstring{I\textsubscript{131}}{I131} & 0.00136 $\pm 0.00047$ & 8.023 d \cite{I131halflife} \\
\texorpdfstring{Te\textsubscript{132}}{Te132} & 1.61 $\pm 0.37$ & 3.180 d \cite{Te132} \\
\texorpdfstring{I\textsubscript{132}}{I132} & 0.017 $\pm 0.11$  & 2.295 h \cite{Te132} \\
\texorpdfstring{Te\textsubscript{133}}{Te133} & 3.93 $\pm 0.15$ & 12.5 m \cite{iaeafissionyield} \\
\texorpdfstring{I\textsubscript{133}}{I133} & 0.153 $\pm 0.053$ & 20.87 h \cite{I133} \\
\texorpdfstring{Te\textsubscript{134}}{Te134} & 6.00 $\pm 0.47$ & 41.8 m \cite{iaeafissionyield} \\
\texorpdfstring{I\textsubscript{134}}{I134} & 0.900 $\pm 0.08$  & 52.5 m \cite{Sonzogni2004} \\
\texorpdfstring{Te\textsubscript{135}}{Te135} & 3.683 $\pm 0.56$ & 19.0 s \cite{iaeafissionyield} \\
\texorpdfstring{I\textsubscript{135}}{I135} & 2.550 $\pm 0.54$ & 6.58 h \cite{Singh2008} \\ \hline
\end{tabular}
\end{center}
\end{table}
